---
title: "sara4r Vignette"
author: "Rafael Hernandez Guzman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sara4r Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

- [Introduction](#intro)
- [How to install](#install)
- [Data preparation](#data)
- [References](#ref)



## Introduction {#intro}

[`sara4r`](http://hydro-geomatic-lab.com/sara4r.html) is an easy way to calculate the rainfall-runoff relation using the Natural Resources Conservation Service - Curve Number method (NRCS-CN method) but includes modifications by Hawkins et al., (2002) about the Initial Abstraction. This graphical user interface follows the programming logic of a previously published software [CN-Idris, Hernández-Guzmán et al., 2011 - CN-Idris: An Idrisi tool for generating curve number maps and estimating direct runoff. Environmental Modelling & Software, 26(12), 1764-1766](https://doi.org/10.1016/j.envsoft.2011.07.006), a raster-based GIS tool that outputs runoff estimates from Land use/land cover and hydrologic soil group maps.
This package is under development at the Institute about Natural Resources Research (INIRENA) from the Universidad Michoacana de San Nicolás de Hidalgo and represents a collaborative effort between the [Hydro-Geomatic Lab](http://hydro-geomatic-lab.com/) (INIRENA) with the Environmental Management Lab (CIAD, A.C.).

## How to install {#install}

`sara4r` package is a Graphical User Interface developed in tcltk and depends on other libraries to run (raster, sp, rgdal). Thus, to make available `sara4r` in the R environment you must install `tcltk2` first, then the `raster`, `sp` and `rgdal` packages.

### First time

First at all, ensure that you have intalled tcltk and tcltk2. 

```{r}
# Load the tcltk package
library(tcltk)

```
Load required packages.

```{r}
# Load the following packages.
library(tcltk2)
library(raster)
library(rgdal)
library(sp)
library(sara4r)
```

Finally, to run our package, just type:
```{r}
sara4r()
```


## Data preparation {#data}
As you can see in the Menu - Help, there are the instructions to make all required files to use `sara4r`. If you go to the installed folder (sara4r\extdata), you will find the `HowtoMake_CNindexFile.xlsx` with all the instructions. In overall terms:

    Land use and land cover map should be reclassified as follow:
    LULC in the first place should be reclassified as 10
    LULC in second place should be reclassified as 20
    ... and so on.
      As an example, Landuse file would be:
        10 Tropical dry forest
        20 Agriculture
        30 Mangrove
        40 Grassland
        50 Evergreen forest
        ...
        
  While the Hydrologic Soil Group map should be reclassified as follow:
    
    HSG A should be reclassified as 1
    HSG B should be reclassified as 2
    HSG C should be reclassified as 3
    HSG D should be reclassified as 4
    Thus, GSH map would be:
        1 A
        2 B
        3 C
        4 D

Finally,

    The logic behind the method is:"
    Landsoil is produced as the sum of LANDUSE and HSG maps. Thus, the possible values it can take are:"
      11 = Landuse 1 (reclassified as 10) with HSG A (reclassified as 1)
      12 = Landuse 1 (reclassified as 10) with HSG B (reclassified as 2)
      13 = Landuse 1 (reclassified as 10) with HSG C (reclassified as 3)
      14 = Landuse 1 (reclassified as 10) with HSG D (reclassified as 4)
      21 = Landuse 2 (reclassified as 20) with HSG A (reclassified as 1)
      22 = Landuse 2 (reclassified as 20) with HSG B (reclassified as 2)
      23 = Landuse 2 (reclassified as 20) with HSG C (reclassified as 3)
      24 = Landuse 2 (reclassified as 20) with HSG D (reclassified as 4)
      31 = Landuse 3 (reclassified as 30) with HSG A (reclassified as 1)
      ...
    Thus, the CN index file would be (csv file):"
      11,12,CNvalue1
      12,13,CNvalue2
      13,14,CNvalue3
      14,15,CNvalue4
      21,22,CNvalue5
      22,23,CNvalue6
      23,24,CNvalue7
      24,25,CNvalue8
      31,32,CNvalue9
      ...
      ...
      ...


## References: {#ref}

> [CN-Idris, Hernández-Guzmán et al., 2011 - CN-Idris: An Idrisi tool for generating curve number maps and estimating direct runoff. Environmental Modelling & Software, 26(12), 1764-1766](https://doi.org/10.1016/j.envsoft.2011.07.006)

> [SARA, Hernández-Guzmán and Ruiz-Luna, 2013. SARA – An enhanced curve number-based tool
for estimating direct runoff. Journal of Hydroinformatics, 15(3), 881-887](https://doi.org/10.2166/hydro.2013.145)

> [sara4r, Hernández-Guzmán et al., 2021. Sara4r: an R graphical user interface (GUI) to estimate watershed surface runoff applying the NRCS – curve number method. Journal of Hydroinformatics, 23(1), 76-87](https://doi.org/10.2166/hydro.2020.087)
